---
title: "Matching"
author: "Ayu Paramudita"
date: "2025-03-21"
output:
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Introduction

**Purpose**

- Learn string matching methods
- **Hopefully**, implement various methods of string distance for industry directory matching


## String Distance

**Comparing two strings to determine how similar or different they are**

- Edit-based (`osa`, `levensthein`)
- Jaro-Winkler
- q-Gram-based (`cosine`, `qgram`)
- Phonetic


## R Package

```{r}

# Load Package
library(stringdist)

```

`amatch(`

  `x,       # what to match`

  `table,   # reference source`

  `method,  # type of string distance method`

  `maxDist, # maximum distance`

  `...)`


## Exact Match

```{r}

# Directory KL
dir_kl <- c("auy", "aypua", "nabi la", "lani", "rani", "rany")

# Directory DSI
dir_dsi <- c("lani", "rani", "ayu", "nabila")

# Exact match
i <- match(dir_kl, dir_dsi)

# Result
data.frame(dir_kl, matched = dir_dsi[i])

```


## Approximate String Matching : Edit-based 

**`osa`** and **`levensthein`**

- Case      : typo
- `maxDist` : max number of typos
- `osa`     : deletion, insertion, substitution and transposition (swapped characters)
- `lv`      : transposition is ignored


```{r}
# osa vs lv
osa1 <- amatch(dir_kl, dir_dsi, method = "osa", maxDist = 1)

lv1 <- amatch(dir_kl, dir_dsi, method = "lv", maxDist = 1)

# Result
data.frame(dir_kl, 
           matched       = dir_dsi[i], 
           amatched_osa1 = dir_dsi[osa1],
           amatched_lv1  = dir_dsi[lv1]
           )

```


## Approximate string matching : Jaro-Winkler

**`jaro-winkler`**

- Case : similarity of name/address
- `maxDist` between 0 and 1
- Higher `maxDist` indicates more dissimilarities


```{r}

# jw
jw <- amatch(dir_kl, dir_dsi, method = "jw", maxDist = 0.2)

# Result
data.frame(dir_kl,
           matched       = dir_dsi[i], 
           amatched_osa1 = dir_dsi[osa1],
           amatched_lv1  = dir_dsi[lv1],
           amatched_jw   = dir_dsi[jw]
           )

```

## Approximate string matching : q-Gram-based

q-Gram-based **`cosine`**

- Case : typo word order
- Splitting words into some parts
- Recognizing words with different order
- Parameter q is how many character split, e.g. q=2 lani (la, an, ni)
- `maxDist` between 0 and 1
- Higher `maxDist` indicates more dissimilarities


```{r}

# cosine
cos <- amatch(dir_kl, dir_dsi, method = "cosine", maxDist = 0.3, q = 2)

# Result
data.frame(dir_kl,
           matched       = dir_dsi[i], 
           amatched_osa1 = dir_dsi[osa1],
           amatched_lv1  = dir_dsi[lv1],
           amatched_jw   = dir_dsi[jw],
           amatched_cos  = dir_dsi[cos]
           )

```

## Thank you!
